---
title: "Case Study 1"
author: "John Partee"
date: "February 22, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```
#Breweries and Beer, an analysis.
I was provided a dataset of breweries, and a dataset of beer, which will be analyzed to answer questions asked at the tail end of this document. In this document, I will address how this data was cleaned up, and analyzed, before presenting answers.

##Import the data:
```{r import}
beers <- read.csv("Beers.csv")
breweries <- read.csv("Breweries.csv")
```

##First, how many breweries are in each state?
```{r perState}
table(breweries$State)
```

##Merge the data by brewery ID, so we can move on with analysis.
```{r merge}
beersAndBreweries <- merge(beers, breweries, by.x = 'Brewery_id', by.y = 'Brew_ID')
#"Names" will be present in both, here is how I fixed that.
dfNames <- names(beersAndBreweries)
#Looking at the original data, col 2 was names for beers, and col 2 was names in breweries. 2+6 (-1 for brew_ID, which disappeared in the merge) = 8
dfNames[2] <- "Beer_Names"
dfNames[8] <- "Brewery_Names"
names(beersAndBreweries) <- dfNames
```

##Ensure it merged right:
```{r checkIt}
head(beersAndBreweries, 6)
tail(beersAndBreweries, 6)
```
The merge was evidently successful, and the columns have useful names.

##There were a number of missing observations in this data, as reported below:
```{r nas}
apply(beersAndBreweries, 2, function(x){sum(is.na(x))})
```
The IBU's and ABV's are the two columns missing values.

##What is the median and mean ABV and IBU per state?
To do this, I'll use dplyr, which has better support for grouping 
```{r groupwiseMedian}
library(dplyr)
groupedbnb <- group_by(beersAndBreweries, State)
medAbv <- summarize(groupedbnb, medABV = median(ABV, na.rm = TRUE))
medIBU <- summarize(groupedbnb, medIBU = median(IBU, na.rm = TRUE))

library(ggplot2)
ggplot(medAbv, aes(x = State, y = medABV)) +
    geom_bar(stat = "identity", fill = "#5EA9BE") +
    coord_cartesian(ylim = range(medAbv$medABV, na.rm = TRUE)) +
    theme(panel.background = element_rect(fill = "#FDF8E2")) +
    ggtitle("Median ABV By State") +
    ylab("Median ABV")

ggplot(medIBU, aes(x = State, y = medIBU)) +
    geom_bar(stat = "identity", fill = "#F3BFB3") +
    coord_cartesian(ylim = range(medIBU$medIBU, na.rm = TRUE)) +
    theme(panel.background = element_rect(fill = "#FDF8E2")) +
    ggtitle("Median IBU By State") +
    ylab("Median IBU")

```






