---
title: "Case Study 1"
author: "Team Greco-Roman"
date: "February 22, 2019"
output: 
  html_document:
    keep_md: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.width = 11, fig.height = 7, fig.align = "center")
```
# Breweries and Beer, an analysis.
I was provided a dataset of breweries, and a dataset of beer, which will be analyzed to answer questions asked at the tail end of this document. In this document, I will address how this data was cleaned up, and analyzed, before presenting answers.

## Import the data:
```{r import}
beers <- read.csv("data/Beers.csv")
breweries <- read.csv("data/Breweries.csv")
```

## First, how many breweries are in each state?
```{r perState}
table(breweries$State)
```

## Merge the data by brewery ID, so we can move on with analysis.
```{r merge}
beersAndBreweries <- merge(beers, breweries, by.x = 'Brewery_id', by.y = 'Brew_ID')
#"Names" will be present in both, here is how I fixed that.
dfNames <- names(beersAndBreweries)
#Looking at the original data, col 2 was names for beers, and col 2 was names in breweries. 2+6 (-1 for brew_ID, which disappeared in the merge) = 8
dfNames[2] <- "Beer_Names"
dfNames[8] <- "Brewery_Names"
names(beersAndBreweries) <- dfNames
```

## Ensure it merged right:
```{r checkIt}
head(beersAndBreweries, 6)
tail(beersAndBreweries, 6)
```
The merge was evidently successful, and the columns have useful names.

## There were a number of missing observations in this data, as reported below:
```{r nas}
colSums(is.na(beersAndBreweries))
```
The IBU's and ABV's are the two columns missing values.

## What is the median and mean ABV and IBU per state?
To do this, I'll use dplyr, which has better support for grouping, and ggplot2 for visualization.
```{r groupwiseMedian}
#Group it all!
library(dplyr)
groupedbnb <- group_by(beersAndBreweries, State)
#get statistics by state
medAbv <- summarize(groupedbnb, medABV = median(ABV, na.rm = TRUE))
medIBU <- summarize(groupedbnb, medIBU = median(IBU, na.rm = TRUE))
#plot it
library(ggplot2)
ggplot(medAbv, aes(x = State, y = medABV)) +
    geom_bar(stat = "identity", fill = "#5EA9BE") +
    theme(panel.background = element_rect(fill = "#FDF8E2"),
          axis.text.x = element_text(angle = 90, hjust = 1),
          plot.title = element_text(hjust = 0.5)) +
    labs(title = "Median ABV By State", y = "Median ABV")

ggplot(medIBU, aes(x = State, y = medIBU)) +
    geom_bar(stat = "identity", fill = "#F3BFB3") +
    theme(panel.background = element_rect(fill = "#FDF8E2"),
          axis.text.x = element_text(angle = 90, hjust = 1),
          plot.title = element_text(hjust = 0.5)) +
    labs(title = "Median IBU By State", y = "Median IBU")
```

## Which state has the most alcoholic beer?
```{r mostABV}
#The highest ABV is:
highestABV <- max(beersAndBreweries$ABV, na.rm = TRUE)
highestABV
#The state with the highest ABV beer is:
as.character(beersAndBreweries$State[which(beersAndBreweries$ABV == highestABV)])
```
Colorado, with an ABV of 0.128.

## Which state has the most bitter (Highest IBU) beer?
```{r mostIBU}
#The highest ABV is:
highestIBU <- max(beersAndBreweries$IBU, na.rm = TRUE)
highestIBU
#The state with the highest ABV beer is:
as.character(beersAndBreweries$State[which(beersAndBreweries$IBU == highestIBU)])
```
Oregon, with an IBU of 138.

## Some further statistics on ABV:
```{r ABVSumStats}
summary(beersAndBreweries$ABV)
#This is easier to see in a histogram:
ggplot(beersAndBreweries, aes(x = ABV)) +
    geom_histogram(fill = "#5EA9BE", color = "#FFFFFF") +
    theme(panel.background = element_rect(fill = "#FDF8E2"))
```
A histogram is a better way to see all of this.

## Is there a correlation between IBU and ABV?
The client asked for a scatter plot, but a linear model and correlation coefficient helps here too!
```{r correlateABVIBU}
ggplot(beersAndBreweries, aes(x = ABV, y = IBU)) + 
    geom_jitter(color = "#F3BFB3") +
    geom_smooth(method = "lm", color = "#5EA9BE")  +
    theme(panel.background = element_rect(fill = "#FDF8E2"),
          plot.title = element_text(hjust = 0.5)) +
    labs(title = "IBU vs ABV")
cor(beersAndBreweries$ABV, beersAndBreweries$IBU, use = "complete.obs")
summary(lm(IBU ~ ABV, data = beersAndBreweries))
```
As we can see, the Pearson correlation coefficient between ABV and IBU is 0.67, indicating a moderate correlation. We fit a simple linear regression model to the data. The model that describes the relationship between ABV and IBU is

> IBU = -34.099 + 1282.037*ABV

We obtained an R² value of 0.4497, indicating that the model explains about 45% of the variability of IBU. The plot also visually suggests that generally the higher the alcohol per volume, the more bitter a beer is.


